---
title: "Tools for GitHub Interaction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tools for GitHub Interaction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(pipfun)
```

# Introduction

The `{pipfun}` package provides a set of tools for interacting with GitHub content, especially designed for working with PIP repositories. These functions act as lightweight wrappers around the `{gh}` package, offering a simpler and more focused interface for specific GitHub tasks related to PIP.


# Managing Credentials

The `{pipfun}` package utilizes the `{gitcreds}` package to manage, store, and authenticate GitHub credentials. `{gitcreds}` provides a straightforward way to securely store your GitHub Personal Access Token (PAT) and ensure seamless access when using `{pipfun}`.

To set up your credentials, you can follow the instructions provided on the [official `{gitcreds}` documentation page](https://gitcreds.r-lib.org/), or refer to Jenny Bryan's excellent guide, [Happy Git With R](https://happygitwithr.com/https-pat.html#gitcreds-package). Both resources provide detailed explanations on how to properly configure your Git credentials.

Please note that `{pipfun}` requires valid GitHub credentials to interact with repositories. If your credentials are not set up correctly with `{gitcreds}`, you will be unable to use `{pipfun}` for any GitHub-related functionality.

# Getting Data from GitHub

The functions in `{pipfun}` are organized to build upon one another, allowing for progressively more complex interactions with GitHub. Below, we'll start by introducing the foundational, lower-level functions that provide core metadata and file information. We will then move towards the higher-level, more abstracted functions that make complex tasks easier.

## Retrieving File Metadata from GitHub

Every file stored in a GitHub repository contains a rich set of metadata, which can be very useful for tracking changes, validating updates, and managing file versions. For example, each file has a unique hash code that serves as a signature for its version. Whenever a file is updated, its hash changes, making it possible to determine whether a file has been modified since the last time it was accessed. This hash is also used to confirm overwriting files in a GitHub repository.

To retrieve this metadata, you can use the `get_file_info_from_gh()` function as follows:

```{r}
ghf_info <- get_file_info_from_gh(
  owner     = getOption("pipfun.ghowner"),
  repo      = "pip_info",
  file_path = "releases.csv",
  branch    = "releases"
)

names(ghf_info)
```

In this example, we use the function to get information about the file `"releases.csv"` in the `"releases"` branch of the `"pip_info"` repository. The `owner` argument specifies the GitHub repository owner, which by default will be the value returned by `getOption("pipfun.ghowner")`. This default value simplifies repeated calls, especially when consistently working within a specific organization's repositories.

The function returns metadata about the file, such as its current hash and other useful information that can be leveraged for version control or ensuring file consistency.

## Retrieving branch metadata from Github
In a similar way, users can use the `get_branch_info_from_gh()` function to retrieve detailed metadata about a specific branch within a GitHub repository. This function provides essential information, such as the branch's name, the latest commit's hash, and details about branch protection rules (if any). Like other {pipfun} functions, it uses the GitHub API for these queries, leveraging your credentials to authenticate requests.
The output can be used for tasks such as verifying the latest changes, ensuring compliance with branch protection rules, or tracking updates across branches in the repository.

```{r}
gh_branch_info <- get_branch_info_from_gh(
  owner     = getOption("pipfun.ghowner"),
  repo      = "aux_ppp",
  branch    = "DEV"
)

names(gh_branch_info)
```

## Downloading a File from GitHub and Reading from Local Drive

Downloading and reading files from GitHub can be challenging due to the diversity of file formats, many of which cannot be read directly. This is where the `get_file_from_gh()` function in `{pipfun}` comes into play.

The `get_file_from_gh()` function simplifies this process by identifying the format of the file, downloading it as a temporary file on your local drive (when necessary), and then loading it directly into memory. This function supports the most common file formats used in the PIP project, including:

-   CSV (`.csv`)
-   Excel files (`.xls`, `.xlsx`)
-   Stata (`.dta`)
-   Quick Serialization (`.qs`)
-   Fast serialization (`.fst`)
-   YAML (`.yaml`)
-   R Data (`.rds`)
-   JSON (`.json`)

This flexibility allows users to work seamlessly with various data types commonly encountered in PIP, without worrying about the specific steps needed to handle each format.

```{r}
dt  <- get_file_from_gh(owner     = getOption("pipfun.ghowner"),
                repo      = "pip_info",
                file_path = "releases.csv",
                branch    = "releases")

head(dt)

```

# Saving Data to GitHub

To save data to GitHub using `{pipfun}`, you need to have appropriate permissions configured in your R environment. Ensure that the token you use with `{gitcreds}` has the necessary permissions to write files to GitHub repositories.

The process of saving data with the `save_to_gh()` function differs from the traditional Git workflow of staging, committing, and pushing changes. Instead, `save_to_gh()` interacts directly with the GitHub API, allowing you to upload data without manually handling Git commands. This approach provides a convenient way to programmatically save or update files in a repository.

The `save_to_gh()` function provides a streamlined way to save or update files in a GitHub repository. If the file already exists, it will be updated with new content. If it does not exist, a new file will be created. 
By default, `save_to_gh()` saves the data in CSV format, but you can choose any of the supported formats mentioned earlier.

Below are examples demonstrating how to use it:

### Example 1: create a new file 

This will create a new file named data_example.csv in the DEV branch of the aux_test repository.

```{r try}
#library(pipfun)

# Example data frame
# df <- data.frame(a = 1:5, b = letters[1:5])
# 
# # Save the data to a repository
# save_to_gh(
#   df       = df,
#   owner    = getOption("pipfun.ghowner"),
#   repo     = "aux_test", 
#   filename = "data_example", 
#   ext      = "csv"
# )

```

### Example 2: updating an existing file 

```{r eval = FALSE}

# Updated data frame
df <- data.frame(a = 6:10, b = letters[6:10])

# Update the file in the repository
save_to_gh(
  df       = df,
  repo     = "aux_test", 
  filename = "data_example", 
  ext      = "csv"
)

```
### Example 3: Saving the Same Data to the Same File (No Change)

```{r eval = FALSE}
# Save the same data to the same file
result_no_change <- save_to_gh(
  df = df,
  repo = "aux_test",
  filename = "data_example",
  ext = "csv"
)

# Check if the data was changed
print(result_no_change$data_change)  # Should be FALSE as the content is identical

```


Additionally, recall that if you have metadata for an existing file, you can pass it directly to the function through the `metadata` argument.

### Understanding its output

The `save_to_gh()` function returns a list invisibly, containing information about the upload or update operation. The key elements of this output list are:

- **`content`**: Metadata about the uploaded or updated file, including its SHA hash, path in the repository, and other details.
- **`commit`**: Information about the commit associated with the upload or update operation, such as the commit SHA and message.
- **`init`**: Metadata of the file before the operation. If the file did not exist, this will be `NULL`.
- **`owner`**: The GitHub username or organization that owns the repository.
- **`repo`**: The name of the GitHub repository where the file was uploaded or updated.
- **`branch`**: The branch of the repository where the file was uploaded or updated.
- **`data_change`**: A logical value indicating whether the file's content was updated (`TRUE`) or remained unchanged (`FALSE`).  
  *Note*: If the file did not exist before the operation, `data_change` will be `TRUE` because creating a new file is considered a change to the repository's data state.



# Delete from Github



# Branches management 

Blah
