---
title: "Using pipfun's Logging System"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using pipfun's Logging System}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#devtools::load_all(".")
library(pipfun)
```

## Intro

The `pipfun` logging system provides a powerful, flexible way to track messages, metadata, function arguments, and error traces across your R codebase — especially when building packages.

Logs are stored in memory and can be written to disk as `.qs` files for persistence.

Use this system to:

- Log errors, warnings, and informative messages
- Capture arguments, return values, and internal states
- Filter, summarize, and debug across modules or packages
- Persist logs between sessions or across users

## Quick Start

```{r}
log_name <- "mylog"
log_init(log_name)

log_info("This is an info message", name = log_name)
log_warn("Something looks odd", name = log_name)
log_error("An error occurred", name = log_name)
```

```{r}
log_has_errors(log_name)
log_summary(log_name)
log_filter(log_name, event = "error")
```

## Logging Concepts

Each log entry is a row in a `data.table`, with fields such as:

- `time`: timestamp
- `event`: one of "info", "warning", or "error"
- `message`: custom message
- `fun`, `package`: where it was logged from
- `args`: captured arguments
- `output`: result (optional)
- `trace`: traceback object

## Creating and Managing Logs

### Initialize a log

```{r}
log_init("devlog")
```

### Add entries

Use the helpers for most use cases:

```{r}
log_info("Process started", name = "devlog", logmeta = list(step = 1))
log_warn("Unexpected format", name = "devlog", logmeta = list(column = "status"))
log_error("Failed to compute results", name = "devlog", logmeta = list(file = "input.csv"))
```

## Filter and Query

```{r}
log_filter(name = "devlog", event = "error")
log_filter(name = "devlog", fun = "my_function")
log_filter(name = "devlog", after = Sys.Date() - 1)
```

## Using Logs in Package Development

`pipfun` is especially helpful when building packages.

### 1. Initialize at load

In your package’s `zzz.R`:

```r
.onLoad <- function(libname, pkgname) {
  pipfun::log_init("pipapi_log")
}
```

### 2. Use in functions

```{r}

# I init the log here, but you can do it in zzz.R, or somewhere else.
log_init("pipapi_log", overwrite = TRUE)

get_data <- function(path, clean = TRUE) {
  log_info("Starting data load", name = "pipapi_log",
           logmeta = list(path = path, clean = clean))

  data <- tryCatch({
    read.csv(path)
  }, error = function(e) {
    log_error("Failed to read CSV", name = "pipapi_log",
              logmeta = list(error = e$message))
    stop(e)
  })

  log_info("Data loaded", name = "pipapi_log",
           logmeta = list(n = nrow(data)))
  return(data)
}
```

### 3. Use with tryCatch for robust pipelines

```{r}
run_step <- function(step_name, expr) {
  tryCatch({
    result <- expr()
    log_info("Step completed", 
             name = "pipeline_log", 
             logmeta = list(step = step_name))
    result
  }, error = function(e) {
    log_error("Step failed", 
              name = "pipeline_log", 
              logmeta = list(step = step_name, error = e$message))
    NULL
  })
}
```

### 4. Enable or Disable Logging Dynamically

You can design your functions to respect a `log = TRUE` argument:

```{r}
foo <- function(x, y, log = TRUE) {
  if (log) {
    log_info("Starting foo()",
             logmeta = list(x = x, y = y))
  }
  result <- x + y
  if (log) {
    log_info("Finished foo()",
             logmeta = list(result = result))
  }
  result
}
```

This makes it easy to disable logging in tests or benchmarks:

```{r}
foo(3, 4, log = FALSE)
```

### 5. Check logs programmatically

```{r}
if (log_has_errors("pipapi_log")) {
  cli::cli_alert_danger("Errors detected!")
  print(log_has_errors("pipapi_log", show = TRUE))
}
```

### 6. Save and Reload Logs

```{r}
tf <- tempfile(fileext = ".qs")
log_save("pipapi_log", path = tf)
log_reset("pipapi_log")
log_load(path = tf, name = "pipapi_log", overwrite = TRUE)
```


## Mastering `log_add()`: The Workhorse of the Logging System

The `log_add()` function is the **core engine** behind the logging system in `pipfun`. While `log_info()`, `log_warn()`, and `log_error()` are convenient wrappers, `log_add()` is designed for **full transparency, flexibility, and programmatic control**.

It allows you to:

- Manually or automatically capture arguments from any function
- Include custom metadata
- Track outputs and computation results
- Register messages across multiple logs
- Define error levels or trace calls

This is the function you use when developing custom pipelines, internal packages, or diagnostics that need deep traceability.

### Function  Structure

```{r, eval=FALSE}
log_add(
  event,
  message,
  name    = getOption("pipfun.log.default"),
  args    = NULL,
  logmeta = NULL,
  output  = NULL,
  .trace  = NULL,
  .env    = rlang::caller_env()
)
```


### `event` — What kind of message is this?

This is the **type** of log entry: it defines how the log is categorized. The most common values are:

- `"info"` – For successful operations or checkpoints
- `"warning"` – When something may be problematic
- `"error"` – When something failed or needs attention

You can also pass custom strings (e.g., `"debug"`, `"data"`) if your logging structure supports it.

**Examples:**

```{r}
lname <- "event_log"
log_init(lname)
log_add("info", "Pipeline started", name = lname)
log_add("warning", "Missing values found", name =lname)
log_add("error", "Failed to merge datasets", name = lname)
```


### `message` — What happened?

A short, **descriptive message** of the event. This field will be shown in logs and summaries, so it should be concise and meaningful.

Best practice: keep messages short but informative — they complement the structured metadata.

**Examples:**

```{r}
log_add("info", "Reading household data", name = lname)
log_add("data", "Poverty line is missing", name = lname)
```


### `name` — Which log to write to?

Each log is a uniquely named object stored in a dedicated environment `.piplogenv`. The `name` argument tells `log_add()` which log to append to.

If not provided, it defaults to the option set in `getOption("pipfun.log.default")`.

**Examples:**

```{r}
# Default log name (from options)
log_add("info", "Starting validation")

# Custom log name
log_init("step3_log")
log_add("info", "Custom step complete", name = "step3_log")
```

This allows multiple logs to coexist (e.g., one per module or process), and even be saved/loaded independently.

### `args` — What were the function arguments?

If provided, this should be a named list of arguments. If omitted (`args = NULL`), then `log_add()` automatically captures the arguments from the calling environment.

This lets you store:
- Inputs used
- Function parameters
- Execution settings

**Examples:**

```{r}
# Explicitly pass arguments
log_add("info", "Manual log", args = list(country = "COL", year = 2022))

# Auto-capture from current function
calculate_stats <- function(x, method = "mean") {
  log_add("info", "summary of x")
  summary(x)
}
calculate_stats(rnorm(100))

log_get()[]
```

If you need to log metadata not available in the function’s arguments, use `logmeta` (see below).

### `logmeta` — Attach extra metadata

`logmeta` is a named list that gets **merged with `args`** and lets you attach metadata that might not be passed as a function argument.

Use it to attach:
- Custom tags
- Status flags
- Source IDs
- Anything else relevant to the context

**Examples:**

```{r}
log_add("info", "Post-processing started",
        logmeta = list(step = "adjustment", dataset = "povcalnet"))

log_add("warning", "Missing PPP detected",
        args = list(country = "ZAF"),
        logmeta = list(stage = "pricing", source = "national_stats"))

log_get()[event == "warning"][.N]
```

This is particularly useful when logging **from within a tryCatch() handler** or from wrappers that don’t have access to the original arguments directly.


### `output` — What did the function return?

You can attach any kind of **output or result** to the log. This makes the log entry a complete record of the input, processing, and result.

This is useful for:

- Storing return values
- Inspecting intermediate results
- Comparing expected vs. actual outputs

**Examples:**

```{r}
result <- stats::lm(mpg ~ wt, data = mtcars)
log_add("model", "Model finished", output = result)
log_filter(event = "model")
```

If the result is large, you might log a summary or just a pointer (e.g., file path or row count):

```{r}
log_add("mtcars", "mtcars summary", 
        output = list(n = nrow(mtcars), 
                      vars = names(mtcars)))
log_filter(event = "mtcars")
log_get()[event == "mtcars"][.N][, output]
```


### `.trace` — What call triggered this?

By default, `log_add()` captures the **immediate call that triggered the log** using `sys.call(-1)`.

You can override it by passing a custom expression or trace object:

```{r}
log_add("lm", "weird reg", .trace = quote(stats::lm(am ~ wt, data = mtcars)))
log_get()[event == "lm"][, trace]
```

This is useful when logging from deeply nested code and you want to reflect **the origin of the issue**, not just the logging wrapper.

### `.env` — Where to look for arguments

This is the environment from which to **capture arguments** when `args = NULL`. It defaults to the function that called `log_add()` (using `rlang::caller_env()`), which is usually correct.

You only need to use this if you're capturing from a different environment on purpose.

**Example:**

```{r, eval=FALSE}
caller <- function(x, y) {
  inner()
}
inner <- function() {
  log_add("caller", "From caller", .env = parent.frame())
}
caller(1, 2)
log_get()[event == "caller"][.N][, args]
```

Without `.env = parent.frame()`, `inner()` would log nothing useful — this makes it possible to step one level up and get what we need.



### Basic use — no `.env` needed

```{r}
log_init("mylog", overwrite = TRUE)

myfun <- function(x = 1, y = 2) {
  log_add("info", "Called myfun", name = "mylog")
}

myfun(5, 10)
log_get("mylog")$args[[1]]
```

This works perfectly: `x` and `y` are captured automatically. No need to set `.env`.

###  Problem: wrapped function logs its own frame

```{r}
wrapped <- function(a = 3) {
  log_add("info", "From wrapped", name = "mylog")
}

outer <- function() {
  a <- 42
  wrapped()
}

log_init("mylog", overwrite = TRUE)
outer()
log_get("mylog")$args[[1]]
```

Oops! The logged arguments come from `wrapped()`, not `outer()` — the values may be defaulted or wrong.


### Solution: pass `.env = parent.frame()`

```{r, eval=FALSE}
wrapped <- function(a = 3) {
  log_add("info", "From wrapped", name = "mylog", .env = parent.frame())
}

outer <- function() {
  a <- 42
  wrapped()
}

log_init("mylog", overwrite = TRUE)
outer()
log_get("mylog")$args[[1]]
```

> Now the arguments from `outer()` are logged correctly — including `a = 42`.

### Best Practices

| Scenario                        | Do you need to set `.env`?          |
|---------------------------------|-------------------------------------|
| Calling `log_add()` directly    | ❌ No (it auto-captures from caller) |
| Calling from a wrapper          | ✅ Yes, set `.env = parent.frame()`  |
| Nested pipelines or metaprogramming | ✅ Yes, be explicit                |
| Building custom logging tools   | ✅ Always set `.env` for consistency |

Using `.env` might look like a technical detail, but it makes your logs **predictable and robust**, especially in large pipelines or packages that delegate work across functions.

This creates a log entry with detailed context that you control entirely.


### Key Takeaways

- `log_add()` is designed to be smart and flexible.
- Use it when you need more control than the helpers offer.
- Combine `args`, `logmeta`, and `output` to produce rich logs.
- Use `.env = environment()` to automatically capture local values.
- Use in conjunction with `tryCatch()` to log recoverable errors.

Let your team know: if they understand `log_add()`, they understand the whole logging system.


## Best Practices for Developers

- Initialize logs early (e.g., in `.onLoad()` or on first use)
- Use structured helpers instead of raw messages. However, I highly recommend you learn how function `log_add()` works and use it as needed. 
- Prefer named `logmeta` fields to store arguments or flags
- Inspect logs interactively or export them for review
- Add `log = TRUE` in your functions for user control

## What’s Next?

Planned improvements to the logging framework include:

- Color-coded `print.piplog()` for interactive sessions
- Log levels (debug, info, warning, error, critical)
- Export to JSON, Markdown, or CSV for diagnostics
- Timeline visualizations of log activity
- `log_args()` and `log_with()` helpers to simplify metadata
- Integration with `rlang::caller_args()` and `current_env()` for richer introspection
