---
title: "Using pipfun's Logging System"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using pipfun's Logging System}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all(".")
```

## Overview

The `pipfun` logging system provides a powerful, flexible way to track messages, metadata, function arguments, and error traces across your R codebase — especially when building packages.

Logs are stored in memory and can be written to disk as `.qs` files for persistence.

Use this system to:

- Log errors, warnings, and informative messages
- Capture arguments, return values, and internal states
- Filter, summarize, and debug across modules or packages
- Persist logs between sessions or across users

## Quick Start

```{r}
log_name <- "mylog"
log_init(log_name)

log_info("This is an info message", name = log_name)
log_warn("Something looks odd", name = log_name)
log_error("An error occurred", name = log_name)
```

```{r}
log_has_errors(log_name)
log_summary(log_name)
log_filter(log_name, event = "error")
```

## Logging Concepts

Each log entry is a row in a `data.table`, with fields such as:

- `time`: timestamp
- `event`: one of "info", "warning", or "error"
- `message`: custom message
- `fun`, `package`: where it was logged from
- `args`: captured arguments
- `output`: result (optional)
- `trace`: traceback object

## Creating and Managing Logs

### Initialize a log

```{r}
log_init("devlog")
```

### Add entries

Use the helpers for most use cases:

```{r}
log_info("Process started", name = "devlog", logmeta = list(step = 1))
log_warn("Unexpected format", name = "devlog", logmeta = list(column = "status"))
log_error("Failed to compute results", name = "devlog", logmeta = list(file = "input.csv"))
```

## Filter and Query

```{r}
log_filter(name = "devlog", event = "error")
log_filter(name = "devlog", fun = "my_function")
log_filter(name = "devlog", after = Sys.Date() - 1)
```

## Using Logs in Package Development

`pipfun` is especially helpful when building packages.

### 1. Initialize at load

In your package’s `zzz.R`:

```r
.onLoad <- function(libname, pkgname) {
  pipfun::log_init("pipapi_log")
}
```

### 2. Use in functions

```{r}
log_init("pipapi_log", overwrite = TRUE)

get_data <- function(path, clean = TRUE) {
  log_info("Starting data load", name = "pipapi_log",
           logmeta = list(path = path, clean = clean))

  data <- tryCatch({
    read.csv(path)
  }, error = function(e) {
    log_error("Failed to read CSV", name = "pipapi_log",
              logmeta = list(error = e$message))
    stop(e)
  })

  log_info("Data loaded", name = "pipapi_log",
           logmeta = list(n = nrow(data)))
  return(data)
}
```

### 3. Use with tryCatch for robust pipelines

```{r}
run_step <- function(step_name, expr) {
  tryCatch({
    result <- expr()
    log_info("Step completed", name = "pipeline_log", logmeta = list(step = step_name))
    result
  }, error = function(e) {
    log_error("Step failed", name = "pipeline_log", logmeta = list(step = step_name, error = e$message))
    NULL
  })
}
```

### 4. Enable or Disable Logging Dynamically

You can design your functions to respect a `log = TRUE` argument:

```r
foo <- function(x, y, log = TRUE) {
  if (log) log_info("Starting foo()", logmeta = list(x = x, y = y))
  result <- x + y
  if (log) log_info("Finished foo()", logmeta = list(result = result))
  result
}
```

This makes it easy to disable logging in tests or benchmarks:

```r
foo(3, 4, log = FALSE)
```

### 5. Check logs programmatically

```{r}
if (log_has_errors("pipapi_log")) {
  cli::cli_alert_danger("Errors detected!")
  print(log_has_errors("pipapi_log", show = TRUE))
}
```

### 6. Save and Reload Logs

```{r}
tf <- tempfile(fileext = ".qs")
log_save("pipapi_log", path = tf)
log_reset("pipapi_log")
log_load(path = tf, name = "pipapi_log", overwrite = TRUE)
```

## Best Practices for Developers

- Initialize logs early (e.g., in `.onLoad()` or on first use)
- Use structured helpers instead of raw messages or warnings
- Prefer named `logmeta` fields to store arguments or flags
- Inspect logs interactively or export them for review
- Add `log = TRUE` in your APIs for user control

## Summarizing Logs

```{r}
log_summary("mylog")
log_summary("mylog", by = c("fun", "event"))
```

## Saving and Reloading

```{r}
log_save("mylog", path = "mylog.qs")
log_reset("mylog")
log_load(path = "mylog.qs", name = "mylog")
```

## Checking for Errors

```{r}
if (log_has_errors("mylog")) {
  cli::cli_alert_danger("Errors found")
}

log_has_errors("mylog", show = TRUE)
```

## What’s Next?

Planned improvements to the logging framework include:

- Color-coded `print.piplog()` for interactive sessions
- Log levels (debug, info, warning, error, critical)
- Export to JSON, Markdown, or CSV for diagnostics
- Timeline visualizations of log activity
- `log_args()` and `log_with()` helpers to simplify metadata
- Integration with `rlang::caller_args()` and `current_env()` for richer introspection
