---
title: "Using pipfun's Logging System"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using pipfun's Logging System}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# Load the package (dev version during development)
devtools::load_all(".")
#library(pipfun)
```


## Overview

The `pipfun` logging system provides a powerful way to track events, diagnostics, and metadata **across functions and packages**. Logs are stored in memory and can be saved to disk as compressed `.qs` files.

Use this system to:
- Track function calls and arguments
- Log errors, warnings, and info messages
- Store outputs and traces
- Filter, inspect, and summarize logs



##  Quick Start

```{r}

log_name <- "mylog" 
log_init(log_name)

# log_info("This is an info message", a = 1:5)
log_info("This is an info message", name = log_name)
# log_warn("Something looks odd", issue = "formatting")
log_warn("Something looks odd", name = log_name)
# log_error("An error occurred", critical = TRUE)
log_error("An error occurred", name = log_name)
```

```{r}
log_has_errors(log_name)
log_summary(log_name)
log_filter(log_name, event = "error")
```

## Logging Concepts

Each log entry is a structured row in a `data.table` with:
- `time`: timestamp
- `event`: "info", "warning", or "error"
- `message`: your custom text
- `fun`, `package`: where it was called
- `args`, `output`, `trace`: full context


## Creating and Managing Logs

### Initialize a new log

```{r}
log_init("devlog")
```

This creates a log stored in `.piplogenv`, and ready to accept entries.

### Add log entries

Use helper functions:

```{r}
log_info("Everything looks fine", id = 123)
log_warn("Something seems off", reason = "missing value")
log_error("Calculation failed", data_id = "XYZ")
```

Each captures arguments and context automatically.


##  Query and Filter Logs

### Show all logs with errors:

```{r}
log_filter(event = "error")
```

### Show logs for a specific function:

```{r}
log_filter(fun = "some_function")
```

### Show logs after a given date:

```{r}
log_filter(after = Sys.Date() - 1)
```


## How to Use Logging Inside Other Packages

You can use `pipfun` logging functions directly in any function of another package you’re developing. Here's a recommended pattern:

### 1. Initialize the log at package load (optional)

In your other package’s `zzz.R`:

```{r , eval=FALSE}
.onLoad <- function(libname, pkgname) {
  pipfun::log_init("pipapi_log")
}
```

### 2. Add logging to individual functions

```{r}
# Inside your pip package function
get_data <- function(path, clean = TRUE) {
  log_info("Starting data load", name = "pipapi_log", path = path, clean = clean)
  
  data <- tryCatch({
    read.csv(path)
  }, error = function(e) {
    log_error("Failed to read CSV", name = "pipapi_log", error = e$message)
    stop(e)
  })

  log_info("Successfully loaded data", name = "pipapi_log", n_rows = nrow(data))
  return(data)
}
```

### 3. Check logs after execution

```{r}
# In your diagnostics script or shiny app
if (log_has_errors("pipapi_log")) {
  cli::cli_alert_danger("Errors detected in pipapi!")
  print(log_has_errors("pipapi_log", show = TRUE))
}
```

### 4. Save logs between sessions

```{r}
tf <- tempfile(fileext = ".qs")
log_save("pipapi_log", path = tf)
```

### 5. Restore logs later

```{r}
log_load(path = tf, name = "pipapi_log")
```


### 6. Notes for Package Developers

- Each log is just a `data.table` stored in memory (`.piplogenv`) — cheap to use and fast to update.
- Use `log_info()`, `log_warn()`, `log_error()` instead of ad hoc `message()` or `warning()` calls.
- You can use the same log across many functions, or different logs per module/function if you prefer.


## Misc 

### Summarize Logs

```{r}
log_summary()
log_summary(by = c("fun", "event"))
```

This gives a quick count of messages per function or type.

### Save and Reload Logs

```{r}
log_save("mylog", path = "mylog.qs")
log_reset("mylog")

log_load(path = "mylog.qs", name = "mylog")
```

You can also load logs from other packages and continue appending to them.


### Check for Errors

```{r}
if (log_has_errors("mylog")) {
  cli::cli_alert_danger("Please fix logged errors before proceeding.")
}
```

You can also view them:

```{r}
log_has_errors("mylog", show = TRUE)
```


### Tips

- Logs are invisible by default.
- Use `print(log_get("mylog"))` or future `print.piplog()` to inspect them.
- Store logs in `.qs` format for speed and compactness.
- Use in `tryCatch()` blocks to log recoverable errors.



## 📚 What’s Next?

Coming soon:
- Print methods with color-coded formatting
- Log levels and thresholds
- Debugging exports (JSON / Markdown / CSV)
- Visual summaries (e.g., timeline charts)
