---
title: "Using pipfun's Logging System"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using pipfun's Logging System}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all(".")
```

## Intro

The `pipfun` logging system provides a powerful, flexible way to track messages, metadata, function arguments, and error traces across your R codebase — especially when building packages.

Logs are stored in memory and can be written to disk as `.qs` files for persistence.

Use this system to:

- Log errors, warnings, and informative messages
- Capture arguments, return values, and internal states
- Filter, summarize, and debug across modules or packages
- Persist logs between sessions or across users

## Quick Start

```{r}
log_name <- "mylog"
log_init(log_name)

log_info("This is an info message", name = log_name)
log_warn("Something looks odd", name = log_name)
log_error("An error occurred", name = log_name)
```

```{r}
log_has_errors(log_name)
log_summary(log_name)
log_filter(log_name, event = "error")
```

## Logging Concepts

Each log entry is a row in a `data.table`, with fields such as:

- `time`: timestamp
- `event`: one of "info", "warning", or "error"
- `message`: custom message
- `fun`, `package`: where it was logged from
- `args`: captured arguments
- `output`: result (optional)
- `trace`: traceback object

## Creating and Managing Logs

### Initialize a log

```{r}
log_init("devlog")
```

### Add entries

Use the helpers for most use cases:

```{r}
log_info("Process started", name = "devlog", logmeta = list(step = 1))
log_warn("Unexpected format", name = "devlog", logmeta = list(column = "status"))
log_error("Failed to compute results", name = "devlog", logmeta = list(file = "input.csv"))
```

## Filter and Query

```{r}
log_filter(name = "devlog", event = "error")
log_filter(name = "devlog", fun = "my_function")
log_filter(name = "devlog", after = Sys.Date() - 1)
```

## Using Logs in Package Development

`pipfun` is especially helpful when building packages.

### 1. Initialize at load

In your package’s `zzz.R`:

```r
.onLoad <- function(libname, pkgname) {
  pipfun::log_init("pipapi_log")
}
```

### 2. Use in functions

```{r}

# I init the log here, but you can do it in zzz.R, or somewhere else.
log_init("pipapi_log", overwrite = TRUE)

get_data <- function(path, clean = TRUE) {
  log_info("Starting data load", name = "pipapi_log",
           logmeta = list(path = path, clean = clean))

  data <- tryCatch({
    read.csv(path)
  }, error = function(e) {
    log_error("Failed to read CSV", name = "pipapi_log",
              logmeta = list(error = e$message))
    stop(e)
  })

  log_info("Data loaded", name = "pipapi_log",
           logmeta = list(n = nrow(data)))
  return(data)
}
```

### 3. Use with tryCatch for robust pipelines

```{r}
run_step <- function(step_name, expr) {
  tryCatch({
    result <- expr()
    log_info("Step completed", 
             name = "pipeline_log", 
             logmeta = list(step = step_name))
    result
  }, error = function(e) {
    log_error("Step failed", 
              name = "pipeline_log", 
              logmeta = list(step = step_name, error = e$message))
    NULL
  })
}
```

### 4. Enable or Disable Logging Dynamically

You can design your functions to respect a `log = TRUE` argument:

```{r}
foo <- function(x, y, log = TRUE) {
  if (log) {
    log_info("Starting foo()",
             logmeta = list(x = x, y = y))
  }
  result <- x + y
  if (log) {
    log_info("Finished foo()",
             logmeta = list(result = result))
  }
  result
}
```

This makes it easy to disable logging in tests or benchmarks:

```{r}
foo(3, 4, log = FALSE)
```

### 5. Check logs programmatically

```{r}
if (log_has_errors("pipapi_log")) {
  cli::cli_alert_danger("Errors detected!")
  print(log_has_errors("pipapi_log", show = TRUE))
}
```

### 6. Save and Reload Logs

```{r}
tf <- tempfile(fileext = ".qs")
log_save("pipapi_log", path = tf)
log_reset("pipapi_log")
log_load(path = tf, name = "pipapi_log", overwrite = TRUE)
```


## Mastering `log_add()`: The Workhorse of the Logging System

While `log_info()`, `log_warn()`, and `log_error()` are convenient, `log_add()` is the engine behind them. Understanding `log_add()` gives you full control over logging and allows you to write more customized, powerful diagnostics.

### the structure

```{r, eval=FALSE}
log_add(
  event,        # required: "info", "warning", "error"
  message,      # required: string
  name    = getOption("pipfun.log.default"),
  args    = NULL,
  logmeta = NULL,
  output  = NULL,
  .trace  = NULL,
  .env    = parent.frame()
)
```

#### Overview of Arguments

| Argument   | Purpose |
|------------|---------|
| `event`    | The type of log entry. Must be `"info"`, `"warning"`, or `"error"`. |
| `message`  | A brief, descriptive message. This is what will appear first when reading the log. |
| `name`     | The name of the log object (defaults to `"default"` or the global option `pipfun.log.default`). |
| `args`     | A named list of arguments to store. If `NULL`, the function automatically captures all local variables from `.env`, including `...`. |
| `logmeta`  | A named list of *additional metadata*. This is merged with `args` and can be used to include tags, flags, file names, IDs, etc. |
| `output`   | An object to attach as the result of the operation. Useful to store the return value, model object, parsed data, etc. |
| `.trace`   | A captured call or stack trace. If `NULL`, defaults to `sys.call(-1)` automatically. |
| `.env`     | The environment from which to capture arguments. Defaults to the calling environment. |


#### Full manual logging

```{r}
log_init(name = "mylog", overwrite = TRUE)
log_add(
  event   = "info",
  message = "Imported user data",
  name    = "mylog",
  args    = list(path = "data/users.csv", format = "csv"),
  output  = "success",
  logmeta = list(stage = "load", user = "analyst")
)
```



## Understanding `.env` in `log_add()`

The `.env` argument in `log_add()` lets you control **which environment's arguments are captured** when recording a log entry. In most simple cases, you don’t need to specify `.env` manually — it defaults to `parent.frame()` and captures the caller's arguments. But in more advanced scenarios, especially when logging from **wrapper functions**, **nested calls**, or **custom logging infrastructure**, setting `.env` explicitly ensures correctness.

Let’s look at some examples.

### Basic use — no `.env` needed

```{r}
log_init("mylog", overwrite = TRUE)

myfun <- function(x = 1, y = 2) {
  log_add("info", "Called myfun", name = "mylog")
}

myfun(5, 10)
log_get("mylog")$args[[1]]
```

> This works perfectly: `x` and `y` are captured automatically. No need to set `.env`.

###  Problem: wrapped function logs its own frame

```{r}
wrapped <- function(a = 3) {
  log_add("info", "From wrapped", name = "mylog")
}

outer <- function() {
  a <- 42
  wrapped()
}

log_init("mylog", overwrite = TRUE)
outer()
log_get("mylog")$args[[1]]
```

> Oops! The logged arguments come from `wrapped()`, not `outer()` — the values may be defaulted or wrong.


### Solution: pass `.env = parent.frame()`

```{r}
wrapped <- function(a = 3) {
  log_add("info", "From wrapped", name = "mylog", .env = parent.frame())
}

outer <- function() {
  a <- 42
  wrapped()
}

log_init("mylog", overwrite = TRUE)
outer()
log_get("mylog")$args[[1]]
```

> ✅ Now the arguments from `outer()` are logged correctly — including `a = 42`.

### Best Practices

| Scenario                        | Do you need to set `.env`?          |
|---------------------------------|-------------------------------------|
| Calling `log_add()` directly    | ❌ No (it auto-captures from caller) |
| Calling from a wrapper          | ✅ Yes, set `.env = parent.frame()`  |
| Nested pipelines or metaprogramming | ✅ Yes, be explicit                |
| Building custom logging tools   | ✅ Always set `.env` for consistency |

Using `.env` might look like a technical detail, but it makes your logs **predictable and robust**, especially in large pipelines or packages that delegate work across functions.

This creates a log entry with detailed context that you control entirely.


### Example 3: Adding logmeta

Use `logmeta` when you want to *augment* the captured arguments:

```r
log_add("warning", "Missing values found",
        name = "mylog",
        logmeta = list(n_missing = 14, variable = "income"))
```

This keeps your logs semantically rich without cluttering the function's local environment.

---

### Example 4: Logging results

```r
model <- lm(mpg ~ wt, data = mtcars)
log_add("info", "Model fit", name = "mylog", output = model)
```

Later, you can inspect or extract the stored model from the log.

---

### Example 5: Logging inside a wrapper

```r
wrapped_call <- function(data, id) {
  log_add("info", "Running wrapped_call", name = "mylog", .env = environment())
  result <- mean(data)
  log_add("info", "Result computed", name = "mylog", output = result)
  result
}
```

This pattern ensures your logs are informative and traceable.

---

### Key Takeaways

- `log_add()` is designed to be smart and flexible.
- Use it when you need more control than the helpers offer.
- Combine `args`, `logmeta`, and `output` to produce rich logs.
- Use `.env = environment()` to automatically capture local values.
- Use in conjunction with `tryCatch()` to log recoverable errors.

Let your team know: if they understand `log_add()`, they understand the whole logging system.


## Best Practices for Developers

- Initialize logs early (e.g., in `.onLoad()` or on first use)
- Use structured helpers instead of raw messages. However, I highly recommend you learn how function `log_add()` works and use it as needed. 
- Prefer named `logmeta` fields to store arguments or flags
- Inspect logs interactively or export them for review
- Add `log = TRUE` in your functions for user control

## What’s Next?

Planned improvements to the logging framework include:

- Color-coded `print.piplog()` for interactive sessions
- Log levels (debug, info, warning, error, critical)
- Export to JSON, Markdown, or CSV for diagnostics
- Timeline visualizations of log activity
- `log_args()` and `log_with()` helpers to simplify metadata
- Integration with `rlang::caller_args()` and `current_env()` for richer introspection
