---
title: "Structured Logging with pipfun"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Structured Logging with pipfun}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
#library(pipfun)
devtools::load_all()
```

# Overview

{pipfun} provides a structured, in-memory logging system designed for package development and data pipelines.

Unlike `message()` or `cat()`, pipfun logs:

* store entries as structured rows

* capture function arguments automatically

* record return values

* attach metadata

* persist to disk with versioning (via {stamp})

Each log is a named piplog object stored in memory and behaves like a data.table.

## The Mental Model

A log is:

* A named table

* Stored in an internal environment

* Appended row-by-row

* Queried later

* Optionally saved to disk

Each log entry records:

| Column   | Meaning                                  |
|----------|------------------------------------------|
| time     | Timestamp                                |
| event    | Event type ("info", "warning", "error", etc.) |
| message  | Human-readable description               |
| fun      | Calling function                         |
| package  | Calling environment                      |
| args     | Captured function arguments              |
| logmeta  | Custom metadata                          |
| output   | Return value or summary                  |
| trace    | Call trace                               |

## 1. Basic Usage
Initialize a Log:

```{r}
log_init("analysis", overwrite = TRUE)

#This creates an empty piplog.

# Add entries
log_add("info", "Pipeline started", name = "analysis")
log_add("warning", "Missing values detected", name = "analysis")
log_add("error", "Model failed to converge", name = "analysis")

# Inspect
log_get("analysis")
```

## 2. Logging Inside Functions
The real power comes from automatic argument capture.
```{r}
compute_mean <- function(x, na_rm = TRUE) {
  
  log_add("info", "Starting computation", name = "analysis")
  
  result <- mean(x, na.rm = na_rm)
  
  log_add("info", "Computation finished",
          name = "analysis",
          output = result)
  
  result
}

compute_mean(c(1, 2, 3, NA))
```

The log now contains:

* the arguments (x, na_rm)
* the return value
* the function call
* the timestamp

Inspect captured arguments:

```{r}
log_get("analysis")$args[[1]]
```

3. Adding Structured Metadata

Use logmeta for domain-specific context.

``` {r}
log_add(
  "validation",
  "Invalid country code",
  name = "analysis",
  logmeta = list(
    field = "country",
    value = "XX",
    stage = "input_check"
  )
)
```

## 4. Filtering logs
Logs are queryable

```{r}
# By event type
log_filter("analysis", event = "error")
# By multiple event types
log_filter("analysis", event = c("warning", "error"))
# By function name
log_filter("analysis", fun = "compute_mean(c(1, 2, 3, NA))")
# By time window
log_filter("analysis", after = Sys.time() - 60)

```

Filtering does not modify the original log.

## 5. Robust Error Logging

A common pattern is structured logging inside tryCatch():

```{r}
log_init("pipeline", overwrite = TRUE)

read_data <- function(path) {
  tryCatch({
    
    log_add("io", "Reading file",
            name = "pipeline",
            logmeta = list(path = path))
    
    data <- read.csv(path)
    
    log_add("io", "File loaded",
            name = "pipeline",
            output = list(rows = nrow(data)))
    
    data
    
  }, error = function(e) {
    
    log_add("error", "File read failed",
            name = "pipeline",
            logmeta = list(
              path = path,
              message = e$message
            ))
    
    NULL
  })
}

```

## 6. Saving and Loading Logs

Logs are stored in memory by default.

To persist them, use `log_save()` and `log_load()` (powered by {stamp}).

#### Save
```{r}
tmp <- tempdir()
stamp::st_init(tmp, alias = "demo")

log_save(
  name  = "analysis",
  id    = file.path(tmp, "analysis_log"),
  alias = "demo"
)
```

#### Load
```{r}
log_reset("analysis")

log_load(
  id    = file.path(tmp, "analysis_log.qs2"),
  name  = "analysis",
  alias = "demo"
)
```

You can also list available versions:
```{r}
log_load(
  id = file.path(tmp, "analysis_log.qs2"),
  version = "available",
  alias = "demo"
)
```

## 7. Recommended Usage Patterns
**Initialize once**

Create logs at:

* package load (.onLoad)

* pipeline start

* test setup

## Log entry and exit
```{r}
log_add("start", "Beginning step", name = "analysis")
# ...
log_add("end", "Step completed", name = "analysis")
```

## Use consistent event types

Common choices:

* "info"

* "warning"

* "error"

* "debug"

* domain-specific tags like "validation", "io", "model"

## Keep messages human-readable

Use logmeta for structured data.

## When Should You Use pipfun Logging?

This system is especially useful when:

* Building R packages

* Writing reproducible pipelines

* Running long batch jobs

* Needing audit trails

* Collaborating across teams

* Saving logs with version control

It is less necessary for:

* Simple scripts

* One-off analyses

* Interactive exploration