---
title: "PIP Release Management"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PIP Release Management}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pipfun)
```


# Intro

The Poverty and Inequality Platform (PIP) by the World Bank, updates Global poverty numbers two times a year around one month before Spring Meetings and Annual Meetings. The `{pipfun}` package allows to manage the PIP releases from your R session. 

# Basics

The most relevant information about PIP is available in the repo *pip_info* of the *PIP-Technical-team* Github group. To get all the releases avialble, run:

```{r}
get_pip_releases()
```

To avoid querying the GitHub API multiple times, `get_pip_releases()` saves the results in the environment `.pipenv`, so it is easily available next time need during the same session.

You can also get the most recent release by running, 

```{r}
get_latest_pip_release()
```


# Create a new PIP release



# Remove a PIP release


# Creation of directories

With each new update, several directories must be created to have the data organized. Everything can be managed by the function `new_pip_release()`, but we provide several other functions to give you finner control. 


