---
title: "PIP Release Management"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PIP Release Management}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all(".")
# library(pipfun)
```


# Introduction

The World Bank's Poverty and Inequality Platform (PIP) updates global poverty data twice a year, typically about a month before the Spring and Annual Meetings. The `{pipfun}` package allows you to manage PIP releases directly from your R session, simplifying tasks related to accessing and managing these updates.

# Basics

The most relevant information for PIP is stored in the *pip_info* repository of the *PIP-Technical-team* GitHub group. To retrieve all available releases, run:

```{r}
old_release <- get_pip_releases()
old_release
```

To minimize GitHub API calls, `get_pip_releases()` saves results in the `.pipenv` environment. This cached information is available for the duration of your session, making it easy to access without repeatedly querying the API.

For the most recent release, simply use:

```{r latest-release}
get_latest_pip_release()
```

# Creating a New PIP Release

To initiate a new release, use the `new_pip_release` function. Although all arguments have default values, it is recommended to specify values explicitly. By default, the release date is set to the current date in `YYYYMMDD` format (e.g., `format(Sys.Date(), "%Y%m%d")`). If the date format is incorrect, an error will occur:

```{r wrong-release, error=TRUE}
new_pip_release(release = "10-28-2020") # incorrect date format
```

The release identity defaults to "PROD" but can be set to "INT" for internal reports or "TEST" for developer testing. Note that "TEST" releases are not intended for transparency or replication purposes. Their main objective is to provide developers of pip a clear environment to experiment with new data and methodologies.

```{r new-pip-release}
tmp_dir <- tempdir()
release_date <- "20241105"
identity     <-  "TEST"

# Create a New PIP Release (NPR)
npr <- new_pip_release(release = release_date,
                       identity = identity,
                       root_dir = tmp_dir)
```

When a new release is created, several steps are performed, all stored in the `new_pip_release()` function's returned object. You can view auxiliary directories set up in the directory specified by the `root_dir` argument. If no directory is specified, `{pipfun}` defaults to an official directory path, provided you have the necessary permissions.

## Working Directories

In this example, we use a temporary directory.

Auxiliary data directory:

```{r aux-dir}
npr$aux_dir |> 
  names() |> 
  fs::dir_exists()
```

Poverty Calculator directory:

```{r pc-dir}
npr$pc_dir |> 
  names() |> 
  fs::dir_exists()
```

## Checking for New Releases

```{r new-release}
release <- get_pip_releases(force = TRUE)
release
```

Here, `force = TRUE` refreshes the release data by calling the GitHub API directly, bypassing the cached data in `.pipenv`. This is useful when you need the latest information. (Note: To disable this function permanently, you may use  option for `{pipfun}`.... (add this option) )

# Removing a PIP Release

If you need to remove a release, perhaps due to a mistake, use `remove_pip_release()`. This function is irreversible, so use it with caution.

```{r remove-release}
removed_release <- 
  remove_pip_release(release     = release_date, 
                     identity    = identity, 
                     working_dir = tmp_dir)

# Check that "20241105-TEST" has been removed
(get_pip_releases(force = TRUE))
```

# Underlying Functions

Both `new_pip_release()` and `remove_pip_release()` act as wrappers around `yyy_aux_dir()` and `yyy_pc_dir()`, where `yyy` can be either `create` or `remove`. You can use these helper functions independently if desired. Additionally, these functions rely on `save_to_gh()`, a wrapper for `gh::gh()` that facilitates easy interactions with GitHub repositories.

Both `new_pip_release()` and `remove_pip_release()` return a list containing all values produced by the underlying helper functions, giving you complete information about the operation's results.

