---
title: "Working with PIP releases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with PIP releases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
devtools::load_all(".")  # Ensure the package is loaded
library(pipfun) 
if (!exists(".pipenv")) {
  message("Initializing .pipenv environment")
  setup_working_release(force = TRUE)  # Ensure .pipenv is set up during knitting
}
```

# Introduction

Creating releases goes beyond organization, information, and vintage control. The main purpose is having a well-defined structure that enables a clear,  efficient and reproducible workflow. 

# Set up a release


```{r setup-release}
(wr <- setup_working_release())

```

Notice that, by default, the working release will correspond to the latest release of the most recent PPP round. In this case `r wr$release` and `r wr$ppp`, respectively. Also, if you try to setup the release again, you'll get an error because there is release already in the `.pipenv` environment.

```{r setup-error, error=TRUE}
(wr <- setup_working_release())

```

If you need to set the environment again, you need to either restart your R session or use argument `force = TRUE`

# The `.pipenv` environment

The `.pipenv` environment contains many pieces of information that are use internally by other functions of the PIP ecosystem. For instance, the `wr` object that is returned invisibly by `setup_working_release()` is available in 

```{r wr-display}
.pipenv$working_release
```

Moreover, all the information contained in the old function `pip_create_globals()` is now available in 

```{r gls-display}
head(.pipenv$gls)
```







